---
title       : Presidential Monarchy & Economic Impacts
subtitle    : Preliminary Proposal for Paper
author      : Kevin Low
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Introduction

The relationship between democracy and economic performance is a complex one that has been studied through time with varying degrees of enthusiasm from academia. Results have been mixed. Even [Francis Fukuyama](http://en.wikipedia.org/wiki/Francis_Fukuyama) who proposed in '[*The End of History and the Last Man*](http://en.wikipedia.org/wiki/The_End_of_History_and_the_Last_Man)' that the Western liberal democracy is the end of the evolution of system of governance backtracked on multiple counts.

We are interested therefore, to study the market responses to such a political shift that reflects a subtle distortion of democratic principles as a step towards a truly autocratic system.

--- 

## Prompt - Kazakhstan

This topic is motivated by the single observation that in 2007, the precedent for a sort of 'presidential monarchy' regime is set down in both Kazakhstan and Uzbekistan, albeit in subtly different conditions:

In the case of Kazakhstan,
> On 18 May 2007, the Parliament of Kazakhstan approved a constitutional amendment which allowed the incumbent president—himself—to run for an unlimited number of five-year terms. This amendment applied specifically and only to Nazarbayev: the original constitution's prescribed maximum of two five-year terms will still apply to all future presidents of Kazakhstan. - [*Wikipedia*](http://en.wikipedia.org/wiki/Nursultan_Nazarbayev)

*Note that Nazarbayev ran for the 2011 election (the first one after the constitutional amendment), which he won.

--- 

## Prompt - Uzbekistan

In the case of Uzbekistan,
> Karimov sought another term in the December 2007 presidential election, despite arguments that he was ineligible due to the two-term limit on the presidency. On November 6, 2007, Karimov accepted the nomination of the Uzbekistan Liberal Democratic Party to run for a third term. On November 19, the Central Election Commission announced the approval of Karimov's candidacy, a decision that Karimov's opponents condemned as illegal. - [*Wikipedia*](http://en.wikipedia.org/wiki/Islam_Karimov)

---

## Market Response Variables

From here on, our job is simply to postulate changes in the behaviours of certain political and economic variables in response to this change.

1. Freedom Indices
 + Economic Freedom Index (Heritage Foundation)
 + EIU Democracy Index
2. Doing Business Ratings
3. Macroeconomic Variable Changes

In each of the above cases, we cannot be sure that there will definitely be a change in response to the prompt. Politically, analysts might already assumed that the presidents would stay in power indefinitely so do not greet those constitutional changes/challenges with surprise. And this could apply in other domains as well.

---

## Freedom Indices



---

## Doing Business


---

## Macroeconomic Variables

Here, we examine the following (data from [WDI](http://data.worldbank.org/data-catalog/world-development-indicators)):

1. GDP Growth
2. FDI as a percentage of GDP
3. Trade as a percentage of GDP

One unfortunate matter is that 2007 happen to be the year of the housing bubble crash in the US, which precipitated the Subprime Mortgage Crisis and resulted in the Great Recession thereafter. The only consolation is that we should not expect the CIS countries to be so plugged into the global economy financially as to suffer. Nevertheless, if we are examining variables such as FDI, Trade and GDP, there is inevitably some degree of impact entering from the global economy. Investors may not be responding to the political change as much as the increase in risk aversion in the financial markets.

```{r silent, echo=FALSE}
# Load necessary libraries
library(dplyr)
library(ggplot2)

# Loading data
simdat <- read.csv("SimpleDat.csv", header = TRUE, nrows = 36)
stem <- c("GDP2005USD", "GDPMKT", "GDPGR", "TRADEPC", "SERVTRADE", "FDIGDPIN", "FDIGDPOUT", "FDIUSDIN", "FDINET")
simdat <- select(simdat, -Series.Name)

# Splitting Datasets
kaz <- simdat[simdat$Country.Name == "Kazakhstan",]
taj <- simdat[simdat$Country.Name == "Tajikistan",]
turk <- simdat[simdat$Country.Name == "Turkmenistan",]
uzb <- simdat[simdat$Country.Name == "Uzbekistan",]

# Extracting Country Names
kaz <- select(kaz, -Country.Name)
uzb <- select(uzb, -Country.Name)
taj <- select(taj, -Country.Name)
turk <- select(turk, -Country.Name)

# Transposing dataframes & then declaring them data frames
kaz <- t(kaz); kaz <- as.data.frame(kaz)
uzb <- t(uzb); uzb <- as.data.frame(uzb)
taj <- t(taj); taj <- as.data.frame(taj)
turk <- t(turk); turk <- as.data.frame(turk)

# Adding variable names
names(kaz) <- stem
names(uzb) <- stem
names(taj) <- stem
names(turk) <- stem

# Generating country variables
kaz <- mutate(kaz, cty = "kaz", year = 1)
taj <- mutate(taj, cty = "taj", year = 1)
uzb <- mutate(uzb, cty = "uzb", year = 1)
turk <- mutate(turk, cty = "turk", year = 1)

# Getting rid of the row names
row.names(kaz) <- NULL
row.names(taj) <- NULL
row.names(uzb) <- NULL
row.names(turk) <- NULL

# Creating the year variables
kaz$year <- seq(1990,2013)
uzb$year <- seq(1990,2013)
taj$year <- seq(1990,2013)
turk$year <- seq(1990,2013)

# Combining the datasets
comdat <- merge(kaz, uzb, all = TRUE)
comdat <- merge(comdat, taj, all = TRUE)
comdat <- merge(comdat, turk, all = TRUE)
```

---

## Macroeconomic Variable: GDP Growth

```{r gdpgrowth, echo=FALSE, warning=FALSE, fig.width = 12}
qplot(data = comdat[which(comdat$cty == "kaz" | comdat$cty == "uzb"),], x = year, geom = "blank") + 
        geom_line(aes(y = GDPGR, col = cty)) + 
        geom_vline(aes(xintercept = 2007), col = "red") + 
        ggtitle("GDP Growth 1990-2013")
```

---

## Macroeconomic Variable: FDI as %GDP

```{r fdi, echo=FALSE, warning=FALSE, fig.width = 12}
qplot(data = comdat[which(comdat$cty == "kaz" | comdat$cty == "uzb"),], x = year, geom = "blank") + 
        geom_line(aes(y = FDIGDPIN, col = cty)) + 
        geom_vline(aes(xintercept = 2007), col = "red") + 
        ggtitle("FDI Inflow 1990-2013")
```


---

## Macroeconomic Variable: Trade as %GDP

```{r trade, echo=FALSE, warning=FALSE, fig.width = 12}
qplot(data = comdat[which(comdat$cty == "kaz" | comdat$cty == "uzb"),], x = year, geom = "blank") + 
        geom_line(aes(y = TRADEPC, col = cty)) + 
        geom_vline(aes(xintercept = 2007), col = "red") + 
        ggtitle("Trade as % GDP 1990-2013")
```

---

## Conclusion

The ultimate goal of the paper is perhaps to ask ourselves: 'Yes, the president is going to continue to rule on. But beyond that, what has changed profoundly that is to affect businesses, economic interactions, opportunities, political stability or other elements that our 'target audience' may care about.