---
title       : Monte Carlo Simulation with R
subtitle    : Slidify Trial Series
author      : Kevin Low
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides

---
## Introduction

# Contents

In this presentation, we are going to perform a Monte Carlo Simulation to study the behaviour of OLS estimated coefficients:

1. Parameter Setup
2. The Model (& running it)
3. The Resulting Distributions

# Loading Libraries

In this simulation, we are going to require ggplot2 library in order to plot the charts we would like to show later, so we begin by loading the necessary library.

```{r libraries}
library(ggplot2)
```


---

## Parameter Setup

We start off by setting seed in order to ensure reproducible 'randomization' and then proceed to define the pre-set parameters of the model we are building

```{r setpar}
set.seed(1234)
a <- 1
b <- 2.5
sigx <- 200
mux <- 7
sigy <- 15
n <- 100
```


---

## The Model

We have the classic OLS model which we embed in a for-loop in order to generate 1000 betas which allows us to build the distribution we want to see.

```{r genbeta}
beta <- NULL
betase <- NULL
for (i in 1:1000) {
        x <- rpois(n, round(runif(1,1,n), digit = 0))
        y <- a + b*x + rnorm(n,0,sigy)
        reg <- summary(lm(y ~ x))
        bet <- reg$coefficients[2] # Extracting the beta for this particular regression
        bse <- reg$coefficients[4] # Extracting the Standard error of beta
        beta <- c(beta, bet)
        betase <- c(betase, bse)
}
estimates <- data.frame(beta, betase, row.names = NULL)
```

---

## Distribution of Beta Coefficients

```{r coeffcharts, fig.height = 4}
ggplot(data = estimates, aes(x = beta)) + 
        geom_histogram(aes(y = ..density..), binwidth = 0.1) + 
        geom_density() + 
        geom_vline(aes(xintercept = b), col = "green") + 
        ggtitle("Distribution of Coefficient Estimates")
```

We see that the mean is indeed on the value `r b` which we have set just now.

---

## Distribution of Standard Errors

```{r secharts, fig.height = 4}
ggplot(data = estimates, aes(x = betase)) + 
        geom_histogram(aes(y = ..density..), binwidth = 0.05) + 
        geom_density() + 
        ggtitle("Distribution of Standard Error of Coefficient Estimates")
```

The standard error appears to follow the chi-sq distribution as we should expect of a normal variable.

---
## Summary

1. We see that the theoretical mean is indeed the monte carlo mean of the distribution of beta
2. We can verify that the resulting betas conforms to a normal distribution
3. We can verify that the standard error of the betas conforms to a chi-sq distribution

